\bf Title here

### 2020-10-01 paper/paper.Rmd
write up for implementation section

```{r toy-interp, fig.height = 8, fig.cap = "\\label{toy-interp} This is a toy example of plotting interpolated points"}
p1 <- holes_2d_better_max_tries %>% explore_trace_interp() + theme(legend.position = "none") + 
  scale_color_botanical(palette = "fern_brown")

p2 <- interrupt_no %>% explore_trace_interp() + theme(legend.position = "none") + 
  scale_color_botanical(palette = "fern_green")

p3 <- holes_2d_better_random %>% explore_trace_interp() + theme(legend.position = "none") + 
  scale_color_botanical(palette = "acacia")

@@

```{r toy-pca, fig.height = 4, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
path1 <- holes_1d_geo$basis %>% 
  flatten_dbl() %>% 
  matrix(ncol = 5, byrow = TRUE)
path2 <- holes_1d_better$basis %>% 
  flatten_dbl() %>% 
  matrix(ncol = 5, byrow = TRUE)
set.seed(1)
sphere <- geozoo::sphere.hollow(p = 5, n = 1000)$point
theoretical <- matrix(c(0, 1, 0, 0, 0,
                        0, -1, 0, 0, 0), nrow = 2, byrow = TRUE)

dt <- rbind(sphere, path1, path2, theoretical)
pca <- dt %>% prcomp()
#summary(pca)
pca_pred <- predict(pca) %>%
  as_tibble() %>%
  mutate(method = c(rep("sphere", nrow(sphere)),
                    holes_1d_geo$method,
                    holes_1d_better$method,
                    rep("theoretical", 2))) %>%
  group_by(method) %>%
  mutate(id = row_number(),
         id = ifelse(method == "sphere", 1, id)) %>%
  ungroup()

p <- pca_pred %>%
  ggplot(aes(x = PC1, y = PC2, color = method))+
  geom_point() +
  geom_point(data = pca_pred %>% filter(method == "theoretical"), 
             size = 10) + 
  geom_point(data = pca_pred %>% filter(method %in% c("search_better", "search_geodesic") & id == 1), size = 10) + 
  theme(aspect.ratio = 1) + 
  scale_color_botanical(palette = "cherry")

@@

```{r fig.cap = "THis is the caption", fig.height=4, fig.align="center"}
# ani <- p + theme(legend.position = "none") +
#   transition_states(id) +
#   shadow_mark()
# 
# animate(ani, nframes = 350, device = "png",
#         renderer = file_renderer("figures/pca/",
#                                  prefix = "pca", overwrite = TRUE))

@@

# Implementation: Ferrn package {#implementation}

1. how does the printed data object looks like?

Do I talk about the work done in the tourr package?

Everything is coded up in a package. Package structure

When the optimisation ends, the data object will be stored and printed (it can be turned off by supplying argument `print = FALSE`). Additional messages during the optimisation can be displayed by argument `verbose = TRUE`.  Notice that the tibble object allows the list-column `basis` to be printed out nicely with the dimension of the projection basis readily available.

We form our first dataset using variables `x1`, `x2`, `x8`, `x9` and `x10`,  run the guided tour with optimiser `search_better` and below shows the first ten rows of the data object.

@@

2.

>>>>

```{r toy-interp, fig.height = 8, fig.cap = "\\label{toy-interp} This is a toy example of plotting interpolated points"}
p1 <- holes_2d_better_max_tries %>% explore_trace_interp() + theme(legend.position = "none") + 
  scale_color_botanical(palette = "fern")

p2 <- interrupt_no %>% explore_trace_interp() + theme(legend.position = "none") + 
  scale_color_botanical(palette = "fern")

p3 <- holes_2d_better_random %>% explore_trace_interp() + theme(legend.position = "none") + 
  scale_color_botanical(palette = "fern")

@@

```{r toy-pca, fig.height = 6, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
# path1 <- holes_1d_geo$basis %>% 
#   flatten_dbl() %>% 
#   matrix(ncol = 5, byrow = TRUE)
# path2 <- holes_1d_better$basis %>% 
#   flatten_dbl() %>% 
#   matrix(ncol = 5, byrow = TRUE)
# set.seed(1)
# sphere <- geozoo::sphere.hollow(p = 5, n = 1000)$point
# theoretical <- matrix(c(0, 1, 0, 0, 0,
#                         0, -1, 0, 0, 0), nrow = 2, byrow = TRUE)
# 
# dt <- rbind(sphere, path1, path2, theoretical)
# pca <- dt %>% prcomp(scale. = TRUE)
# #summary(pca)
# pca_pred <- predict(pca) %>%
#   as_tibble() %>%
#   mutate(method = c(rep("sphere", nrow(sphere)),
#                     holes_1d_geo$method,
#                     holes_1d_better$method,
#                     rep("theoretical", 2))) %>%
#   group_by(method) %>%
#   mutate(id = row_number(),
#          id = ifelse(method == "sphere", 1, id)) %>%
#   ungroup()
# 
# p <- pca_pred %>%
#   ggplot(aes(x = PC1, y = PC2, color = method))+
#   geom_point() +
#   geom_point(data = pca_pred %>% filter(method == "theoretical"), 
#              size = 10) + 
#   geom_point(data = pca_pred %>% filter(method %in% c("search_better", "search_geodesic") & id == 1), size = 10) + 
#   theme(aspect.ratio = 1) + 
#   scale_color_botanical(palette = "cherry")
# 
# p

p <- bind_rows(holes_1d_geo, holes_1d_better) %>% 
  bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5), 
                   index = tourr::holes(), raw_data = data) %>% 
  bind_theoretical(matrix(c(0, -1, 0, 0,0), nrow = 5), 
                   index = tourr::holes(), raw_data  = data)%>%
  explore_space_pca(col = method)  + 
  scale_color_botanical(palette = "cherry") + 
  theme(legend.position = "bottom")

@@

```{r fig.cap = "THis is the caption", fig.height=4, fig.align="center"}
# ani <- p + theme(legend.position = "none") +
#   transition_states(id) +
#   shadow_mark()
# 
# ani <- bind_rows(holes_1d_geo, holes_1d_better) %>% 
#   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5), 
#                    index = tourr::holes(), raw_data = data) %>% 
#   bind_theoretical(matrix(c(0, -1, 0, 0,0), nrow = 5), 
#                    index = tourr::holes(), raw_data  = data)%>%
#   explore_space_pca(col = method, animate = TRUE)  + 
#   scale_color_botanical(palette = "cherry") 
# 
# animate(ani, nframes = 350, device = "png",
#         renderer = file_renderer("figures/pca/",
#                                  prefix = "pca", overwrite = TRUE))

@@

# Implementation {#implementation}

(further notice on the implementation in the tour package will be discussed in the tour developer meeting.)
The implementation of this projection has been divided into two pacakges, where the data collection happens when a tour is running and hence implemented in the existing `tourr` package. The diagnostics of the algorithm via plots and other auxiliary functions have been implemented in a new pacakge, `ferrn`.

When the optimisation ends, the data object will be stored and printed (it can be turned off by supplying argument `print = FALSE`). Additional messages during the optimisation can be displayed by argument `verbose = TRUE`.  Notice that the tibble object allows the list-column `basis` to be printed out nicely with the dimension of the projection basis readily available. Below presents a sample of the data object.

@@

Once the data object has been obtained, the package, `ferrn`, provides automatic diagnostic plots as described in Section \ref{vis-diag}. The structure of package functionality has been outlined in Table \ref{table:ferrn-glossary}. The four main functions plots, `explore_trace_search()`, `explore_trace_interp()`, `explore_space_pca()`, and `explore_space_tour()`, produce the diagnostic plots upon a simple supply of the data objects. Auxiliary functions, `get_*()` and `bind_*()`, provide handy solutions to extract certain components in the data object. They modulate the extraction and binding procedures in the main functions while being convenient to use on their own.

```{r eval = FALSE, echo = TRUE}
# Below are codes that generate the plots in Section 4 
# using functions from ferrn package
bind_rows(holes_2d_better_max_tries %>% mutate(max_tries = 500), 
          holes_2d_better %>% mutate(max_tries = 25)) %>% 
  explore_trace_search(group = max_tries)
```

```{r eval = FALSE, echo = TRUE}
bind_rows(holes_2d_better_max_tries %>% mutate(group = "max_tries: 500"), 
          interrupt_no %>% mutate(group = "no interruption"), 
          holes_2d_better_random %>% mutate(group = "better_random")) %>% 
  mutate(group = fct_relevel(group, c("max_tries: 500", 
                                      "no interruption", 
                                      "better_random"))) %>% 
  explore_trace_interp(group = group) + 
  theme(legend.position = "none")
```

```{r eval = FALSE, echo = TRUE}
bind_rows(holes_1d_geo, holes_1d_better) %>% 
  bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5), 
                   index = tourr::holes(), raw_data = data) %>% 
  bind_theoretical(matrix(c(0, -1, 0, 0,0), nrow = 5), 
                   index = tourr::holes(), raw_data  = data)%>%
  explore_space_pca(col = method)  + 
  scale_color_botanical(palette = "cherry")
```

```{r eval = FALSE, echo = TRUE}
bind_rows(holes_1d_geo, holes_1d_better) %>% 
  explore_space_tour(color = method)
```

## Extract pieces from the data object

 `get_*()` extract and, in some cases, manipulate certain components from the existing data object, for example,  `get_best()` extracts the best basis found by the algorithm and `get_interp()` extracts the observations in the interpolation and prepare them ready to be supplied to `explore_trace_interp()`.

## Bind external information

`bind_*()` incorporate additional information that is outside the tour optimisation. Given a simulated dataset, the most interesting projection is known ahead of time, so does its basis.  `bind_theoretical()` can be utilised to incorporate this best possible basis to the existing data object with the supply of the index function and original data for producing the index value. This binding is used in the space plots to compare the position of the theoretical best basis to the ones being searched by the algorithm. The following chunk presents its usage.

```{r bind-theoretical-demo, eval = FALSE, echo = TRUE}
theo_best <- matrix(c(0, 1, 0, 0, 0), nrow = 5)
holes_1d_better %>% 
  bind_theoretical(matrix = theo_best, index = tourr::holes(), raw_data = data) 
```

`bind_random()` and `bind_random_matrix()` generates a large number of points on the surface of a high dimensional sphere and attaches it to the existing data object. This is also useful in the space plots since it preserves the spherical structure of the spaces of the projection bases, in the computation of PCA for `explore_space_pca()` and animation of `explore_space_tour()`.

## Botanical color palette and reorder factor levels

A collection of color palettes from Australian native plants has been created to match the theme of the package Quantitative palettes, daisy, banksia and cherry, are recommended for comparing between algorithms, while sequantial palettes, fern and acacia are better for trace plots. These color palettes can be accessed via

- `scale_color_botanical(palette = "...")` for ggplot constructions, and

- `explore_space_tour(..., color = botantical_palettes$...)` for a full space exploration

When colored by `info`, users may want to map a particular color to a certain level to emphasize certain elements in the algorithm, `relevel_*()` allows the user to manipulate the level in the `info` column in the data object with the default being the chronological order that each level occurs in the optimisation.

\begin{table}
\centering
\begin{tabular}{|wl{5cm} wl{6cm}|}
\hline
\multicolumn{2}{|l|}{\textbf{Main functions}}\\
\multicolumn{2}{|l|}{explore\_trace\_search()}\\
\multicolumn{2}{|l|}{explore\_trace\_interp()}\\
\multicolumn{2}{|l|}{explore\_space\_pca() }\\
\multicolumn{2}{|l|}{explore\_space\_tour()}\\
& \\
\multicolumn{2}{|l|}{\textbf{Auxiliary functions}}\\
get\_best() & bind\_theoretical() \\
get\_start() & bind\_random() \\
get\_interp() & bind\_random\_matrix() \\
get\_search\_count() & \\
get\_basis\_matrix() & \\
& \\
relevel\_geo() & botanical\_palettes:  fern, acacia\\
relevel\_better() & \hspace{3mm} daisy, banksia, and cherry\\
& botanical\_pal() \\
& scale\_color\_botanical()\\
\hline
\end{tabular}
\caption{A glossary of functions available in ferrn package. The main functions are four automatic plotting functions and auxiliary functions provides easier manipulation on the data object.}
\label{table:ferrn-glossary}
\end{table}

### 2020-10-01 paper/paper.tex
write up for implementation section

\hypertarget{implementation}{%
\section{Implementation: Ferrn package}\label{implementation}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  how does the printed data object looks like?
\end{enumerate}

Do I talk about the work done in the tourr package?

Everything is coded up in a package. Package structure

When the optimisation ends, the data object will be stored and printed
(it can be turned off by supplying argument \texttt{print\ =\ FALSE}).
Additional messages during the optimisation can be displayed by argument
\texttt{verbose\ =\ TRUE}. Notice that the tibble object allows the
list-column \texttt{basis} to be printed out nicely with the dimension
of the projection basis readily available.

We form our first dataset using variables \texttt{x1}, \texttt{x2},
\texttt{x8}, \texttt{x9} and \texttt{x10}, run the guided tour with
optimiser \texttt{search\_better} and below shows the first ten rows of
the data object.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\item
\end{enumerate}

>>>>

\hypertarget{implementation}{%
\section{Implementation}\label{implementation}}

(further notice on the implementation in the tour package will be
discussed in the tour developer meeting.) The implementation of this
projection has been divided into two pacakges, where the data collection
happens when a tour is running and hence implemented in the existing
\texttt{tourr} package. The diagnostics of the algorithm via plots and
other auxiliary functions have been implemented in a new pacakge,
\texttt{ferrn}.

When the optimisation ends, the data object will be stored and printed
(it can be turned off by supplying argument \texttt{print\ =\ FALSE}).
Additional messages during the optimisation can be displayed by argument
\texttt{verbose\ =\ TRUE}. Notice that the tibble object allows the
list-column \texttt{basis} to be printed out nicely with the dimension
of the projection basis readily available. Below presents a sample of
the data object.

@@

Once the data object has been obtained, the package, \texttt{ferrn},
provides automatic diagnostic plots as described in Section
\ref{vis-diag}. The structure of package functionality has been outlined
in Table \ref{table:ferrn-glossary}. The four main functions plots,
\texttt{explore\_trace\_search()}, \texttt{explore\_trace\_interp()},
\texttt{explore\_space\_pca()}, and \texttt{explore\_space\_tour()},
produce the diagnostic plots upon a simple supply of the data objects.
Auxiliary functions, \texttt{get\_*()} and \texttt{bind\_*()}, provide
handy solutions to extract certain components in the data object. They
modulate the extraction and binding procedures in the main functions
while being convenient to use on their own.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Below are codes that generate the plots in Section 4 }
\CommentTok{# using functions from ferrn package}
\KeywordTok{bind_rows}\NormalTok{(holes_2d_better_max_tries }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{max_tries =} \DecValTok{500}\NormalTok{), }
\NormalTok{          holes_2d_better }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{max_tries =} \DecValTok{25}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{explore_trace_search}\NormalTok{(}\DataTypeTok{group =}\NormalTok{ max_tries)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{bind_rows}\NormalTok{(holes_2d_better_max_tries }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{group =} \StringTok{"max_tries: 500"}\NormalTok{), }
\NormalTok{          interrupt_no }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{group =} \StringTok{"no interruption"}\NormalTok{), }
\NormalTok{          holes_2d_better_random }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{group =} \StringTok{"better_random"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{group =} \KeywordTok{fct_relevel}\NormalTok{(group, }\KeywordTok{c}\NormalTok{(}\StringTok{"max_tries: 500"}\NormalTok{, }
                                      \StringTok{"no interruption"}\NormalTok{, }
                                      \StringTok{"better_random"}\NormalTok{))) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{explore_trace_interp}\NormalTok{(}\DataTypeTok{group =}\NormalTok{ group) }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"none"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{bind_rows}\NormalTok{(holes_1d_geo, holes_1d_better) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{bind_theoretical}\NormalTok{(}\KeywordTok{matrix}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{nrow =} \DecValTok{5}\NormalTok{), }
                   \DataTypeTok{index =}\NormalTok{ tourr}\OperatorTok{::}\KeywordTok{holes}\NormalTok{(), }\DataTypeTok{raw_data =}\NormalTok{ data) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{bind_theoretical}\NormalTok{(}\KeywordTok{matrix}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{-1}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{), }\DataTypeTok{nrow =} \DecValTok{5}\NormalTok{), }
                   \DataTypeTok{index =}\NormalTok{ tourr}\OperatorTok{::}\KeywordTok{holes}\NormalTok{(), }\DataTypeTok{raw_data  =}\NormalTok{ data)}\OperatorTok{%>%}
\StringTok{  }\KeywordTok{explore_space_pca}\NormalTok{(}\DataTypeTok{col =}\NormalTok{ method)  }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{scale_color_botanical}\NormalTok{(}\DataTypeTok{palette =} \StringTok{"cherry"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{bind_rows}\NormalTok{(holes_1d_geo, holes_1d_better) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{explore_space_tour}\NormalTok{(}\DataTypeTok{color =}\NormalTok{ method)}
\end{Highlighting}
\end{Shaded}

\hypertarget{extract-pieces-from-the-data-object}{%
\subsection{Extract pieces from the data
object}\label{extract-pieces-from-the-data-object}}

\texttt{get\_*()} extract and, in some cases, manipulate certain
components from the existing data object, for example,
\texttt{get\_best()} extracts the best basis found by the algorithm and
\texttt{get\_interp()} extracts the observations in the interpolation
and prepare them ready to be supplied to
\texttt{explore\_trace\_interp()}.

\hypertarget{bind-external-information}{%
\subsection{Bind external information}\label{bind-external-information}}

\texttt{bind\_*()} incorporate additional information that is outside
the tour optimisation. Given a simulated dataset, the most interesting
projection is known ahead of time, so does its basis.
\texttt{bind\_theoretical()} can be utilised to incorporate this best
possible basis to the existing data object with the supply of the index
function and original data for producing the index value. This binding
is used in the space plots to compare the position of the theoretical
best basis to the ones being searched by the algorithm. The following
chunk presents its usage.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{theo_best <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{nrow =} \DecValTok{5}\NormalTok{)}
\NormalTok{holes_1d_better }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{bind_theoretical}\NormalTok{(}\DataTypeTok{matrix =}\NormalTok{ theo_best, }\DataTypeTok{index =}\NormalTok{ tourr}\OperatorTok{::}\KeywordTok{holes}\NormalTok{(), }\DataTypeTok{raw_data =}\NormalTok{ data) }
\end{Highlighting}
\end{Shaded}

\texttt{bind\_random()} and \texttt{bind\_random\_matrix()} generates a
large number of points on the surface of a high dimensional sphere and
attaches it to the existing data object. This is also useful in the
space plots since it preserves the spherical structure of the spaces of
the projection bases, in the computation of PCA for
\texttt{explore\_space\_pca()} and animation of
\texttt{explore\_space\_tour()}.

\hypertarget{botanical-color-palette-and-reorder-factor-levels}{%
\subsection{Botanical color palette and reorder factor
levels}\label{botanical-color-palette-and-reorder-factor-levels}}

A collection of color palettes from Australian native plants has been
created to match the theme of the package Quantitative palettes, daisy,
banksia and cherry, are recommended for comparing between algorithms,
while sequantial palettes, fern and acacia are better for trace plots.
These color palettes can be accessed via

\begin{itemize}
\item
  \texttt{scale\_color\_botanical(palette\ =\ "...")} for ggplot
  constructions, and
\item
  \texttt{explore\_space\_tour(...,\ color\ =\ botantical\_palettes\$...)}
  for a full space exploration
\end{itemize}

When colored by \texttt{info}, users may want to map a particular color
to a certain level to emphasize certain elements in the algorithm,
\texttt{relevel\_*()} allows the user to manipulate the level in the
\texttt{info} column in the data object with the default being the
chronological order that each level occurs in the optimisation.

\begin{table}
\centering
\begin{tabular}{|wl{5cm} wl{6cm}|}
\hline
\multicolumn{2}{|l|}{\textbf{Main functions}}\\
\multicolumn{2}{|l|}{explore\_trace\_search()}\\
\multicolumn{2}{|l|}{explore\_trace\_interp()}\\
\multicolumn{2}{|l|}{explore\_space\_pca() }\\
\multicolumn{2}{|l|}{explore\_space\_tour()}\\
& \\
\multicolumn{2}{|l|}{\textbf{Auxiliary functions}}\\
get\_best() & bind\_theoretical() \\
get\_start() & bind\_random() \\
get\_interp() & bind\_random\_matrix() \\
get\_search\_count() & \\
get\_basis\_matrix() & \\
& \\
relevel\_geo() & botanical\_palettes:  fern, acacia\\
relevel\_better() & \hspace{3mm} daisy, banksia, and cherry\\
& botanical\_pal() \\
& scale\_color\_botanical()\\
\hline
\end{tabular}
\caption{A glossary of functions available in ferrn package. The main functions are four automatic plotting functions and auxiliary functions provides easier manipulation on the data object.}
\label{table:ferrn-glossary}
\end{table}
